## Handedness in ARIES 

### Sample populations

ARIES time points/groups:
```{r}
aries.times <- c(cord="cord",
                 F7="age 7",
                 "15up"="age 15-17",
                 F24="age 24",
                 antenatal="age 29 (mothers during pregnancy)",
                 FOM="age 47 (mothers)",
                 FOF="age 54 (fathers)",
                 parents="mothers (age 47) and fathers (age 54)")
aries.times <- data.frame(aries.times=names(aries.times), description=aries.times)
kable(aries.times, row.names=F)
```

```{r,results="asis"}
library(tableone)
for (time in names(ewas.ret)) {
    for (model in names(ewas.ret[[time]])) {
        ewas <- ewas.ret[[time]][[model]]
        catvars <- names(which(apply(ewas$design, 2, function(var) length(unique(var)) < 10)))
        strata <- colnames(ewas$design)[grep("handedright", colnames(ewas$design))]
        surrogates <- colnames(ewas$design)[grep("^sv[0-9]+$", colnames(ewas$design))]
        covars <- setdiff(colnames(ewas$design), c("(Intercept)", surrogates, strata)) 
        ret <- CreateTableOne(vars=covars, 
                              strata=strata,
                              data=as.data.frame(ewas$design),
                              factorVars=catvars)
        cat(paste0("\n**", time, " ", model, " model:\nmethylation ~ ", paste(c("handedness", covars), collapse=" + "), "**\n"))
        print(kableone(ret))
    }
}
```

### EWAS results


```{r}
kable(do.call(rbind, lapply(names(ewas.ret), function(time) {
    do.call(rbind, lapply(names(ewas.ret[[time]]), function(model) {
        ewas <- ewas.ret[[time]][[model]]$table
        idx <- which(ewas$p.adjust < 0.05)
        if (length(idx) == 0) return(NULL)
        data.frame(time=time, model=model, ewas[idx,])
    }))
})))
```


Shown below are the single CpG sites with the lowest p-value in each EWAS.
```{r}
ret <- do.call(rbind, lapply(names(ewas.ret), function(time) {
    do.call(rbind, lapply(names(ewas.ret[[time]]), function(model) {
        ewas <- ewas.ret[[time]][[model]]
        idx <- which.min(ewas$table$p.value)
        data.frame(time=time, model=model, site=rownames(ewas$table)[idx], ewas$table[idx,])
    }))
}))
ret$p.value <- format(ret$p.value, scientific=T)
kable(ret,digits=4)
```

Agreement is *not* high between the time points.

P-values for these sites in all time points (basic model):
```{r}
top.sites <- unique(as.character(ret$site))
ret.p <- sapply(top.sites, function(site) {
    sapply(ewas.ret, function(ewas.by.time) ewas.by.time$basic$table[site,"p.value"])
})
ret.p <- format(ret.p,scientific=T,digits=3)
kable(ret.p)
```

... and full model:
```{r}
ret.p <- sapply(top.sites, function(site) {
    sapply(ewas.ret, function(ewas.by.time) ewas.by.time$full$table[site,"p.value"])
})
ret.p <- format(ret.p,scientific=T,digits=3)
kable(ret.p)
```


Correlation of the top 100 effects across all time points
(the top 100 were selected from the model corresponding to the columns in the table).

Basic model
```{r}
for (time in names(ewas.ret))
    ewas.ret[[time]]$basic$top <- with(ewas.ret[[time]]$basic, rownames(table)[order(table$p.value)[1:100]])
ret <- sapply(ewas.ret, function(a) {
    sapply(ewas.ret, function(b) {
        cor(a$basic$table[a$basic$top,"estimate"], b$basic$table[a$basic$top,"estimate"], use="p")
    })
})
kable(ret,digits=3)
```

.. and full model:
```{r}
for (time in names(ewas.ret))
    ewas.ret[[time]]$full$top <- with(ewas.ret[[time]]$full, rownames(table)[order(table$p.value)[1:100]])
ret <- sapply(ewas.ret, function(a) {
    sapply(ewas.ret, function(b) {
        cor(a$full$table[a$full$top,"estimate"], b$full$table[a$full$top,"estimate"], use="p")
    })
})
kable(ret,digits=3)
```


Agreement about the top 100 associations (basic model):
```{r}
ret <- sapply(ewas.ret, function(a) {
    sapply(ewas.ret, function(b) {
        length(intersect(a$basic$top, b$basic$top))
    })
})
kable(ret,digits=3)
```

... and full model:
```{r}
ret <- sapply(ewas.ret, function(a) {
    sapply(ewas.ret, function(b) {
        length(intersect(a$full$top, b$full$top))
    })
})
kable(ret,digits=3)
```

### DMR results

```{r}
kable(do.call(rbind, lapply(names(dmr.ret), function(time) {
    do.call(rbind, lapply(names(dmr.ret[[time]]), function(model) {
        dmrs <- dmr.ret[[time]][[model]]
        idx <- which(dmrs$p.adjust < 0.05 & dmrs$n > 1)
        if (length(idx) == 0) return(NULL)
        data.frame(time=time, model=model, dmrs[idx,c("chr","start","end","estimate","se","p.value","p.adjust")])
    }))
})), row.names=F)
```
